---
title: "Fantasy Football 2015 1 Week Competitions"
author: "Isaac J. Faber"
date: "September 26, 2015"
output: html_document
---

This package provides a set of functions to create optimal line-ups for use in 1 week NFL fantasy football leagues.  The purpose of this narrative is to show how to use the FantasyFootball2015 package.  First, load the packages.
```{r}
library(devtools)
library(stringdist)
install_github('FantasyFootball2015', 'isaacfab')
library(FantasyFootball2015)
```

For starters, you need to read in the salaries from Draft Kings.  You can bring in any competition as long as it is for the NFL and the 2015 season.  This .csv file is provided by the site when you build a line up.  We want to create a descriptive name for each player in order to join each row with other information later.  Here we also create a variable (binary only 1 or 0) that if set to 1 requires that player to be in a given lineup.

```{r}
data<-read.csv("DKSalaries.csv",header=TRUE,strip.white=TRUE)

data$DescriptiveName<-paste(data$Name,data$teamAbbrev,data$Position,sep=" ")

data$Require<-0

print(head(data))
```

Several things should be observed from the raw data.  First, each player has a 'salary' and a line up must be selected such that the total salary does not exceed a given cap (default is 50,000).  In addition, a column has the average number of points for a given player.  This number is the mean for this season.  Draft Kings will include players in the line up selection that are either injured or suspended.  You, obviously, want to avoid selecting these players.  On the actual website their names have a flag next to them but for analysis or automation we have to remove them ourselves.  The first function the FantasyFootball2015 package provides is an automated method for retrieving current injuries or suspensions (from Yahoo).  The first function returns players that are currently ineligible to play.  Keep in mind that these functions scrape current data!  

```{r}
data.inel<-FantasyFootball2015::ineligible()

print(data.inel[1:3,])
```

The second function includes players that are game day calls, such as those that are questionable.  In practice, this column will have the actual status of the player.  

```{r}
data.gdc<-FantasyFootball2015::gamedaycall()

print(data.gdc[1:3,])
```

We want to join the injury information to the draft kings data. 
```{r}
data$injury<-"None"
data$gdc<-"None"
for(i in 1:length(data[,1])){
    j<-stringdist::amatch(data$DescriptiveName[i],data.inel$DescriptiveName,maxDist=5)
    if(!is.na(j)){
      data$injury[i]<-data.inel[["Injury Type"]][j]
    }
}

for(i in 1:length(data[,1])){
    j<-stringdist::amatch(data$DescriptiveName[i],data.gdc$DescriptiveName,maxDist=5)
    if(!is.na(j)){
      data$gdc[i]<-data.gdc[["Status"]][j]
    }
}

print(data[1:5,])

```
Now you have to make some decisions on what players you want to exclude.  Obviously, you want to rule out anyone who is 'Out' but there is also a category of 'Questionable.'  This category often includes high profile players that may need a case-by-case decision.  For this example, we will only keep players that are either not injured or are probable.  You can subset however you like.
```{r}
data<-data[data$gdc=="None"|data$gdc=="P",]
table(data$gdc)
```
Now that you have removed any players that will likely not be on the field, it is time to start thinking about how to forecast player performance.  The following function scrapes fantasy point predictions from 9 fantasy football websites and gives two types of averages.  The only argument that this function accepts is an integer value for the week of the season (must be an integer).  Some of the forecasts require the week and some do not.  Like the injuries, this function only returns current values.
```{r}
forecast<-FantasyFootball2015::CombinedForecasts(8)

print(forecast[1:3,])
```
Need to combine these forecasts with the data.
```{r}
n<-colnames(forecast)
n<-n[2:length(n)]
for(i in 1:length(n)){
  data[[n[i]]]<-0
}

#combine the forecasts and the Draft Kings file
for(i in 1:length(data[,1])){
  j<-stringdist::amatch(data$DescriptiveName[i],forecast$DescriptiveName,maxDist=5)
  if(!is.na(j)){
    for(k in 1:length(n)){
      data[[n[k]]][i]<-forecast[[k+1]][j]
    }
  }
}

```

The quality of the lineup that is played is dependent on how good the forecasts are.  There is room here for you to complete your own forecasts from any number of sources such as regression, machine learning, other expert predictions, ownership rates and so on.  However, even if you have perfect forecasts you would still have the problem of knowing which combination of players results in the greatest amount of fantasy points (the possible combinations of teams is enormous).    

The following function preforms a linear program and returns the optimal lineup from a given set of forecasts.  In other words if your forecast are perfect this function will give you the team with the highest score.  There are three functional arguments; our data, the forecast column name and the salary cap (defaults to 50,000).  Here are three optimal lineups using different sources of forecasting and constraints.

```{r}
#Optimal Lineup with Average Points Per Game forecast
opt<-FantasyFootball2015::DKoptLineUpNFL(data,"AvgPointsPerGame")
print(opt[,1:4])

#Optimal Lineup with ESPN forecast and require that Russell Wilson is in the lineup
data$Require[data$DescriptiveName=="Russell Wilson Sea QB"]<-1
espn_opt<-FantasyFootball2015::DKoptLineUpNFL(data,"ESPN_f")
print(espn_opt[,1:4])

#Optimal Lineup with Average forecast but lower salary cap
avg_opt<-FantasyFootball2015::DKoptLineUpNFL(data,"Average_Robust",45000)
print(avg_opt[,1:4])
```
This next function in the package is handy for collecting information in order to conduct back testing.  The history functions scrape historical by-player game level performance from ESPN or NFL.com for a given week and season.  ESPN offers historical data for the 2015 and 2014 seasons.  However, NFL.com has a much longer window (back to 1970).  

ESPN and NFL.com have different types of data.  ESPN includes all offensive positions excluding kickers and defensive/ST stats.  NFL.com offers all offensive information, including kickers, but nothing on defense.  There still needs to be some amount of calculating to determine the number of fantasy points earned (different for each site).

```{r}
#historical performance for week 5 of the 2014 season
hist<-FantasyFootball2015::historyESPN(5,2014)

print(hist[,1:4])

#historical performance for week 8 of the 1997 season
hist<-FantasyFootball2015::historyNFL(8,1997)

print(hist[,1:4])
```
